\speculake: Hiding Malware in Speculative Execution

### 2018-12-11 paper/implementation.tex
impl pass, nested spec pass

% TODO: this could move elsewhere?
We demonstrate the speed that information can be decrypted via the speculative
world, and we vary the channel width of the side channel from 1 to 12 bits to
measure its performance. At low channel width, reading from the cache side
channel requires timing reads from only 2 locations, while at 12-bits, the side
channel requires reading $2^{12}$ locations. On the other hand, there is a fixed
overhead per speculative iteration that favors increased channel width to
maximize bandwidth. As shown in Figure~\ref{fig:spec_bandwidth}, 8 bits is the
optimal side channel width, allowing us to decrypt over 5,000 bits per second
(625 Bytes/sec).
% Iterations: 20

>>>>

% TODO: this could move elsewhere?
We demonstrate the speed that information can be decrypted via the speculative
world, and we vary the channel width of the side channel from 1 to 12 bits to
measure its performance. At low channel width, reading from the cache side
channel requires timing reads from only 2 locations, while at 12-bits, the side
channel requires reading $2^{12}$ locations. On the other hand, there is a fixed
overhead per speculative iteration that favors increased channel width to
maximize bandwidth. As shown in Figure~\ref{fig:spec_bandwidth}, 8 bits is the
optimal side channel width, allowing us to decrypt over 5,000 bits per second
(625 Bytes/sec). We note an improvement over this rate by loading multiple
values into the probe array during speculation. Instead of a single
probe array of 1024 entries communicating 10-bits, we can split the array into
four sections of 256 entries each, and signal four times (one per section) during
speculation. This provides a 32-bit channel overall, while only having to probe
1024 entries. This method is limited by the limited $\mu$~operation budget,
however, our implementation using these parameters (four sections of 256
entries) is able to decrypt over 11~Kbps (1,425 Bytes/sec).

% Iterations: 20

### 2018-12-11 paper/payloads.tex
impl pass, nested spec pass

We can overcome this by having the trigger program communicate the decryption
key to the payload program via the branch predictor. Previous work has analyzed
the capability of the branch predictor to distribute keys to other
programs~\cite{evtyushkin2016jump, lee2017inferring, aciiccmez2007predicting}
and demonstrated that this is not a secure method of key distribution. However,
the goal of using the branch predictor is not share a single secret via entry
point but rather to use multiple jumps to construct the decryption key
speculatively. To accomplish this, we use multiple speculative entry points,
each that derives a different decryption key before calling a common decryption
routine. Since the exact speculative entry point is determined by the trigger
program, an analyst cannot discover the decryption key directly from the payload
program.

Still, an analyst could enumerate all potential entry points, testing each one
until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
there are (at most) only 1~million entry points, providing just 20~bits of
security, trivial for an analyst to brute force.

To increase security, we instead use nested speculation to \emph{chain} entry
points together. Rather than derive the key from a single entry point, we have
each potential entry point perform another indirect jump that the CPU cannot
immediately resolve, forcing it to speculate while already executing
speculatively. The predicted target of that jump will also be determined by the
trigger program. Meaning that an analyst that is (somehow, possibly via oracle,
or attacks similar to~\cite{aciiccmez2007predicting}) able to determine the
location of first speculative jump is still unable to reconstruct the whole key
as future speculative jumps are determined solely by the trigger program which
the analyst does not have access to.

For example, if the trigger program makes 30 training jumps, followed by 10
additional indirect jumps, and the payload program performs the same 30 training
jumps before a stall, the CPU will predict the payload program will also perform
the next 10 jumps. Speculatively the CPU will attempt to follow each of these
jumps speculatively.

If each jump has the potential to land in up to 1024 possible locations, this
would provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of
security. We describe our implementation of nested speculative execution in
Section~\ref{subsec:impl-aes}, where we achieve 128~bits of security.
Figure~\ref{fig:nested-spec} demonstrates this in practice, the trigger program
running a series of training jumps followed by indirect jumps influences the
payload program to follow the same path. Any analyst attempting to follow this
path without the trigger program would need to correctly jump to each of the 10
addresses to load the correct key. The security gained here is not a result of a
single entry point remaining secret but rather multiple independent jumps by the
unknown trigger program providing numerous landing points for each jump.

>>>>

We can overcome this by having the trigger program communicate the decryption
key to the payload program via the branch predictor. While prior work has used
the branch predictor to exfiltrate keys from other sensitive
processes~\cite{lee2017inferring}, we inject a key into the speculating payload program
from the external trigger program. To do this, we use
multiple speculative entry points, each that derives a unique decryption key
before calling a common decryption
routine. Since the exact speculative entry point is determined by the trigger
program, an analyst cannot trivially discover the decryption key directly from the payload
program.

Still, an analyst could enumerate all potential entry points, testing each one
until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
there are (at most) only 1~million possible entry points, providing just 20~bits of
security, trivial for an analyst to brute force. An analyst simply needs to test
each of the $2^{20}$ entry potential entry points to discover the correct key.

To increase security,
we instead use nested speculation to \emph{chain} entry
points together. Rather than derive the key from a single entry point, we have
each potential entry point perform another indirect jump that the CPU cannot
immediately resolve, forcing it to speculate while already executing
speculatively. In other words, in the speculative world, we make an
indirect jump that depends on a cache-evicted variable, prompting the CPU to double-speculate.
The predicted target of that jump will also be driven by the
trigger program's (mis)training of the indirect branch predictor.
On CPUs that support double (or arbitrarily nested)
speculation, we can repeat this process, with each new subsequent entry point
determined by the trigger program. At each entry point, we shift in additional
bits to a register as the AES key. Without the trigger program, an analyst
cannot determine the path the payload program will take speculatively.

%Meaning that an analyst that is (somehow, possibly via oracle,
%or attacks similar to~\cite{aciiccmez2007predicting}) able to determine the
%location of first speculative jump is still unable to reconstruct the whole key
%as future speculative jumps are determined solely by the trigger program which
%the analyst does not have access to.
%

As an illustrating example,
imagine a trigger program makes 30 training jumps, followed by 10
additional indirect jumps, and the payload program performs the same 30 training
jumps before a stall. At this point,
the CPU will predict the payload program will also perform
the next 10 jumps, speculatively following the pattern of the trigger program.

If each nested speculative jump has the potential to land in 4096 ($2^{12}$ possible locations,
each
entry point can shift in 12-bits to the key, for a total of 120-bits over the
10 jumps before calling the common decryption routine. A key constructed in this
way would be infeasible
for an analyst to brute force, as the payload program yields no information
about which of the potential $2^{120}$ keys will be derived.

We describe our implementation of nested speculative execution in
Section~\ref{subsec:impl-aes}, where we speculatively derive a 128~bit AES key.
Figure~\ref{fig:nested-spec} demonstrates this in practice, the trigger program
running a series of training jumps followed by indirect jumps influences the
payload program to follow the same path.
% Could cut? Redundant with above now
%An analyst attempting to follow this
%path without the trigger program would need to correctly jump to each of the 10
%addresses to load the correct key. The secret key created here is not a result
%of a single entry point remaining secret (via the branch predictor) but rather
%multiple independent jumps by the unknown trigger program providing numerous
%potential speculative entry points for each jump.
